Epidemiological birth-death models with partner notification

### 2024-06-06 article/article.tex
Transmission from notified partners is possible

The equations~\ref{eq:Us}, \ref{eq:p}, \ref{eq:p-n}-\ref{eq:p-b} have the following closed form solutions:
\begin{equation}
\scriptsize
\begin{split}
&\begin{cases}
U(t) = \frac{\lambda + \psi}{2\lambda} +  \frac{c_1}{2\lambda}\Big(\frac{E_1(t) - 1}{E_1(t) + 1}\Big)\\
\\
p^{(i)}(t) = \Big(\frac{E_2(t_i) + 1}{E_1(t) + 1}\Big)^2E_3(t, t_i) \\
\\
p_n^{(i)}(t) =  e^{(\lambda + \psi) (t -t_i)}\\
\\
p_b^{(i)}(t) =  e^{(\lambda + \phi) (t -t_i)}
\\
\\
p_{pb}^{(i)}(t) =  \sqrt{p^{(i)}(t)p_n^{(i)}(t)} \\
\\
p_{pa}^{(i)}(t) =  e^{(\phi-\psi) (t -t_i)} p_{pb}^{(i)}(t) 
\end{cases},\\
& \textit{where } c_1 = \sqrt{(\lambda - \psi)^2 + 4 \lambda\psi\rho},\; c_2 = \frac{c_1 + \lambda - \psi}{c_1 - \lambda + \psi},\\
& ~~~~~~~~E_1(t) = c_2 e^{c_1 (t -T)},\; E_2(t_i) = c_2 e^{c_1 (t_i -T)},\; E_3(t, t_i) = e^{c_1 (t -t_i)}\\
\end{split}\label{eq:ps}
\end{equation}

@@

\begin{equation}
\begin{split}
p_{m(s1,s2)}^{(i)}(t,t_r) &= p_{top=s1}^{(h)}(t)U_p(t_h,t_r)p_{bottom=s2}^{(i)}(t_h)\textit{\color{gray},}\\
&\textit{\color{gray}where $t_h$ is the time of the hidden partner subtree start,}\\
&\textit{\color{gray}~~~~~~~~$t_r$ is the first notification time of the hidden partner.}\\
\end{split}
\end{equation}\label{eq:mixed}

@@

\begin{equation}
p_{bottom=s2}^{(i)}(t_h) = 
\begin{cases}
p_n^{(i)}(t)&\textit{\color{gray} if $i$ is a notifier tip ($s2=n$),}\\
p^{(i)}(t)&\textit{\color{gray} otherwise ($s2=-$, the standard state)}\\
\end{cases}
\end{equation}

@@

We will treat the case when the branch is in the partner state ($s1=p$) in a similar way:
\begin{equation}
\begin{split}
p_{top=p}^{(h)}(t, t_r) &\approx 
\begin{cases}
\frac{p_{pb}^{(h)}(t) -e^{-(\lambda + \psi)(t_{h} - t)}}{U(t_{h})}&\textit{\color{gray} if $t_h \leq t_r$}\\
p_{pb}^{(r)}(t)\frac{p_{pa}^{(h)}(t_r) -e^{-(\lambda + \psi)(t_{h} - t_r)}}{U(t_{h})}&\textit{\color{gray} if $t \leq t_r \leq t_h$}\\
\frac{p_{pa}^{(h)}(t) -e^{-(\lambda + \psi)(t_{h} - t)}}{U(t_{h})}&\textit{\color{gray} if $t > t_r$}\\
\end{cases}\\
&\textit{\color{gray}where $t_r$ is the first notification time of the hidden partner.}\\
\end{split}
\end{equation}

Note that we wrote down equation~(\ref{eq:mixed}) as if the start of the hidden partner tree $t_h$ was known. In practice, it is not the case as we need to integrate over all the possible times between the branch start $t_{parent(i)}$ and its end $t_i)$. To simplify the calculations, we will take the time in the middle of this interval instead:

@@

Finally, we will treat the mixed case with two hidden partners using similar principals:
\begin{equation}
\begin{split}
p_{m(o,-,n)}^{(i)}(t,t_r) &\approx \Big(p_{o}^{(h1)}(t) -e^{-(\lambda + \psi)(t_{h1} - t)}\Big)\frac{U_p(t_{h1},t_{r})}{U(t_{h1})}\cdot
\\
&~~~~\Big(p^{(h2)}(t_{h1}) -e^{-(\lambda + \psi)(t_{h2} - t_{h1})}\Big)\frac{U_p(t_{h2},t_{i})}{U(t_{h2})}\cdot
\\
&~~~~~p_n^{(i)}(t_{h2})\textit{\color{gray},}\\
&\textit{\color{gray}where $t_{h1} \approx t + \frac{1}{3}\Big(min(t_{r}, t_i) - t\Big)$,}\\
&\textit{\color{gray}~~~~~~~~$t_{h2} \approx t + \frac{2}{3}\Big(min(t_{r}, t_i) - t\Big)$,}\\
&\textit{\color{gray}~~~~~~~~$t_{r}$ is the earliest notification time of the }\\
&\textit{\color{gray}~~~~~~~~~~~~hidden partner at the top of the branch (o,-).}\\
\end{split}\label{eq:p_mixed_mixed}
\end{equation}

@@

0 \textit{\color{gray} ~if $t_r < t_j$}\\
\textit{\color{gray} (not possible as once notified partners do not transmit)}

@@

\textit{\color{gray} (oriented (partner) branch evolution followed by standard sampling and no notification)}

@@

~~~+p_{m(o,-)}^{(i)}(t_j,t_r)\psi\rho(1-\upsilon) \textit{\color{gray} ~i.e. branch $i$ contains a hidden partner notified by $r$ at $t_r > t_i$, }\\

@@

 \textit{\color{gray} ~~~(mixed oriented-standard branch evolution followed by sampling and no notification)}\\
 ~~~+p_{m(o,n)}^{(i)}(t_j,t_i)\psi\rho\upsilon \textit{\color{gray} ~i.e. branch $i$ contains a hidden partner notified both by $r$ and $i$}\\
 \textit{\color{gray} ~~~(mixed oriented-notifier branch evolution followed by sampling and notification)}

@@

~~~+p_{m(o,-,n)}^{(i)}(t_j,t_r)\psi\rho\upsilon  \textit{\color{gray} ~i.e. branch $i$ contains a hidden partner notified by $r$ at $t_r > t_i$, }\\

@@

 \textit{\color{gray} ~~~(mixed oriented-standard-notifier branch evolution followed by sampling and notification)}

@@

\textit{\color{gray} (oriented (partner) branch evolution followed by} \\

@@

~~~+p_{m(o,-)}^{(i)}(t_j,t_r)\psi\rho(1-\upsilon) \textit{\color{gray} ~i.e. branch $i$ contains a hidden partner notified by $r$ at $t_j \leq t_r \leq t_i$, }\\

@@

 \textit{\color{gray} ~~~(mixed oriented-standard branch evolution followed by sampling and no notification)}\\
 ~~~+p_{m(o,n)}^{(i)}(t_j,t_r)\psi\rho\upsilon \textit{\color{gray} ~i.e. branch $i$ contains a hidden partner notified both by $r$ and $i$}\\
 \textit{\color{gray} ~~~(mixed oriented-notifier branch evolution followed by sampling and notification)}

@@

~~~+p_{m(o,-,n)}^{(i)}(t_j,t_r)\psi\rho\upsilon  \textit{\color{gray} ~i.e. branch $i$ contains a hidden partner notified by $r$ at $t_j \leq t_r \leq t_i$, }\\

@@

l^{(n,n)}(i,r) &= \begin{cases}0 \textit{\color{gray} ~if $t_r < t_j$}\\
\textit{\color{gray} (not possible as once notified partners do not transmit)}\\

@@

p_n^{(r)}(t_j) p_{pa}^{(i)}(t_r) \phi \upsilon \textit{\color{gray} ~if $i$ got notified at time $t_j \leq t_r \leq t_i$ and notified}\\
\textit{\color{gray} (notifier branch evolution followed by} \\
\textit{\color{gray}~notified partner branch evolution,} \\

@@

In the other case the internal node $i$'s branch starts as a partner who will eventually get notified by a tip $r$.
There are two possibilities: either the node $i$ corresponds to the partner or the partner is hidden along the $i$'s branch and $i$ corresponds to someone within $i$'s recipient subtree (e.g., if $i=C$ in Fig.~\ref{fig:pn-branches}(2a)). In the first case, the $i$'s branch evolution is represented by $p_{pb}^{(i)}(t_j)$ with a transmission event at the end of it, where the donor is known (probability density of $\lambda$): The donor corresponds to the subtree containing the tip representing the sampling of the individual $i$.

In the second case, the $i$'s branch evolution is represented by $p_{m(o,-)}^{(i)}(t_j,t_r)$ with a transmission event at the end of it, where the donor is unknown (probability density of $2\lambda$): any of the subtrees can be the donor.
 %Moreover, this recipient's branch cannot be notified by our partner (as in our model notified partners do not notify further):
Again, we will consider three possibilities for $i$'s child nodes: (2.2.1) when both $i_0$ and $i_1$ are internal nodes; (2.2.2) when both $i_0$ and $i_1$ are tips; and (2.2.3) when one of them is an internal node and the other one is a tip.

In any of these scenarios, the notification must have happened after the partner tip branch start (as one of the model assumptions is that partners do not transmit after notification). Hence:
\begin{equation}
\scriptsize
l^{(n)}(i, r) = 0 \textit{\color{gray}~if $t_i > t_r$} \\
\label{eq:lp-zero}
\end{equation}

@@

\begin{equation}
\scriptsize
\begin{split}
l^{(n)}(i, r) = &p_{pb}^{(i)}(t_j) \lambda \cdot
\Big(~l^{(n)}(i_0,r)l^{(\not{n})}(i_1) \textit{\color{gray}~$\leftarrow$ $i_0$ is notified by $r$} \\
& ~~~~~~~~~~~~~~~+ l^{(\not{n})}(i_0)l^{(n)}(i_1,r)~\Big) \textit{\color{gray}~$\leftarrow$ $i_1$ is notified by $r$} \\
& + p_{m(o,-)}^{(i)}(t_j,t_r) 2 \lambda l^{(\not{n})}(i_0)l^{(\not{n})}(i_1) \textit{\color{gray}~$\leftarrow$ hidden partner}\\
 \end{split}
\label{eq:lp-int-int}
\end{equation}

@@

\begin{equation}
\scriptsize
\begin{split}
l^{(n)}(i, r) = &p_{pb}^{(i)}(t_j) \lambda \cdot
\Big(~l^{(n,\not{n})}(i_0,first(r,i_1))l^{(\not{n},n)}(i_1) \textit{\color{gray}~$\leftarrow$ $r$ and $i_1$ notified $i_0$, $i_0$ did not notify} \\
&~~~~~~~~~~~~~~~+ l^{(n,n)}(i_0,first(r,i_1))l^{(n,n)}(i_1, i_0) \textit{\color{gray}~$\leftarrow$ $r$ and $i_1$ notified $i_0$, $i_0$ notified $i_1$} \\
&~~~~~~~~~~~~~~~+l^{(\not{n},n)}(i_0)l^{(n,\not{n})}(i_1,first(r,i_0)) \textit{\color{gray}~$\leftarrow$ $r$ and $i_0$ notified $i_1$, $i_1$ did not notify} \\
&~~~~~~~~~~~~~~~+ l^{(n,n)}(i_0, i_1)l^{(n,n)}(i_1,first(r,i_0)) \textit{\color{gray}~$\leftarrow$ $r$ and $i_0$ notified $i_1$, $i_1$ notified $i_0$} \\
&~~~~~~~~~~~~~~~+l^{(n,n)}(i_0,r)l^{(n,\not{n})}(i_1, i_0) \textit{\color{gray}~$\leftarrow$ $r$ notified $i_0$, $i_0$ notified $i_1$, $i_1$ did not notify} \\
&~~~~~~~~~~~~~~~+ l^{(n,\not{n})}(i_0, i_1)l^{(n,n)}(i_1,r) \textit{\color{gray}~$\leftarrow$ $r$ notified $i_1$, $i_1$ notified $i_0$, $i_0$ did not notify} \\
&~~~~~~~~~~~~~~~+l^{(n,\not{n})}(i_0,r)l^{(\not{n},\not{n})}(i_1) \textit{\color{gray}~$\leftarrow$ $r$ notified $i_0$, $i_0$ and $i_1$ did not notify} \\
&~~~~~~~~~~~~~~~+ l^{(\not{n},\not{n})}(i_0)l^{(n,\not{n})}(i_1,r)~\Big) \textit{\color{gray}~$\leftarrow$ $r$ notified $i_1$, $i_0$ and $i_1$ did not notify} 
\\
&+ p_{m(o,-)}^{(i)}(t_j,t_r) 2 \lambda \cdot
\Big(~l^{(n,\not{n})}(i_0,i_1)l^{(\not{n},n)}(i_1) \textit{\color{gray}~$\leftarrow$ $r$ notified a hidden partner,} \\
&\textit{\color{gray}~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~$i_1$ notified $i_0$, $i_0$ did not notify} \\
&~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~+ l^{(n,n)}(i_0,i_1)l^{(n,n)}(i_1, i_0)\textit{\color{gray}~$\leftarrow$ $r$ notified a hidden partner,} \\
&\textit{\color{gray}~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~$i_1$ notified $i_0$, $i_0$ notified $i_1$} \\
&~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~+l^{(\not{n},n)}(i_0)l^{(n,\not{n})}(i_1,i_0) \textit{\color{gray}~$\leftarrow$ $r$ notified a hidden partner,} \\
&\textit{\color{gray}~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~$i_0$ notified $i_1$, $i_1$ did not notify} \\
&~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~+l^{(\not{n},\not{n})}(i_0)l^{(\not{n},\not{n})}(i_1)~\Big) \textit{\color{gray}~$\leftarrow$ $r$ notified a hidden partner,} \\
&\textit{\color{gray}~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~$i_0$ and $i_1$ did not notify} \\
\\
&\textit{\color{gray}~where } \color{gray}first(a,b) = 
\begin{cases}
a \textit{~if $t_a \leq t_b$}\\
b \textit{~if $t_a > t_b$.}
\end{cases} 
 \end{split}
\label{eq:lu-tip-tip}
\end{equation}

@@

\begin{equation}
\scriptsize
\begin{split}
l^{(n)}(i, r) = &p_{pb}^{(i)}(t_j) \lambda \cdot
\Big(~l^{(n,n)}(i_0,r)l^{(n)}(i_1,i_0) \textit{\color{gray}~$\leftarrow$ $r$ notified $i_0$, $i_0$  notified $i_1$} \\
&~~~~~~~~~~~~~~~+ l^{(n,\not{n})}(i_0,r)l^{(\not{n})}(i_1) \textit{\color{gray}~$\leftarrow$  $r$ notified $i_0$, $i_0$ did not notify} \\
&~~~~~~~~~~~~~~~+l^{(\not{n}, n)}(i_0)l^{(n)}(i_1,first(r,i_0)) \textit{\color{gray}~$\leftarrow$ both $r$ and $i_0$ notified $i_1$} \\
&~~~~~~~~~~~~~~~+ l^{(\not{n},\not{n})}(i_0)l^{(n)}(i_1,r)~\Big) \textit{\color{gray}~$\leftarrow$ $r$ notified $i_1$, $i_0$ did not notify} 
\\
&+ p_{m(o,-)}^{(i)}(t_j,t_r) 2\lambda \cdot
\Big(~l^{(\not{n},n)}(i_0)l^{(n)}(i_1,i_0) \textit{\color{gray}~$\leftarrow$ $r$ notified a hidden partner,} \\
&\textit{\color{gray}~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~$i_0$ notified $i_1$} \\
&~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~+ l^{(\not{n},\not{n})}(i_0)l^{(\not{n})}(i_1)~\Big) \textit{\color{gray}~$\leftarrow$ $r$ notified a hidden partner,} \\
&\textit{\color{gray}~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ $i_0$ did not notify}\\
\textit{\color{gray}~where } &\color{gray}first(a,b) = 
\begin{cases}
a \textit{~if $t_a \leq t_b$}\\
b \textit{~if $t_a > t_b$.}
\end{cases} 
 \end{split}
\label{eq:lp-tip-int}
\end{equation}

>>>>

The equations~\ref{eq:Us}, \ref{eq:p}, \ref{eq:p-n}-\ref{eq:p-b} have the following closed form solutions:
\begin{equation}
\scriptsize
\begin{split}
&\begin{cases}
U(t) = \frac{\lambda + \psi}{2\lambda} +  \frac{c_1}{2\lambda}\Big(\frac{E(t) - 1}{E(t) + 1}\Big)\\
\\
p^{(i)}(t) = \Big(\frac{E(t_i) + 1}{E(t) + 1}\Big)^2 e^{c_1 (t -t_i)} \\
\\
p_n^{(i)}(t) =  e^{(\lambda + \psi) (t -t_i)}\\
\\
p_b^{(i)}(t) =  e^{(\lambda + \phi) (t -t_i)}
\\
\\
p_{pb}^{(i)}(t) =  \sqrt{p^{(i)}(t)p_n^{(i)}(t)} \\
\\
p_{pa}^{(i)}(t) =  e^{(\phi-\psi) (t -t_i)} p_{pb}^{(i)}(t) 
\end{cases},\\
& \textit{where } c_1 = \sqrt{(\lambda - \psi)^2 + 4 \lambda\psi\rho},\; c_2 = \frac{c_1 + \lambda - \psi}{c_1 - \lambda + \psi},\\
& ~~~~~~~~E(t) = c_2 e^{c_1 (t -T)} \\
\end{split}\label{eq:ps}
\end{equation}

@@

\begin{equation}
\begin{split}
p_{m(s1,s2)}^{(i)}(t,t_r) &= p_{top=s1}^{(h)}(t)U_p(t_h,t_r)p_{bottom=s2}^{(i)}(t_h)\textit{\color{gray},}\\
&\textit{\color{gray}where $t_h$ is the time of the hidden partner subtree start,}\\
&\textit{\color{gray}~~~~~~~~$t_r$ is the first notification time of the hidden partner.}\\
\end{split}\label{eq:p_mixed}
\end{equation}

@@

\begin{equation}
p_{bottom=s2}^{(i)}(t_h) = 
\begin{cases}
p_n^{(i)}(t_h)&\textit{\color{gray} if $i$ is a notifier tip ($s2=n$),}\\
p^{(i)}(t_h)&\textit{\color{gray} otherwise ($s2=-$, the standard state)}\\
\end{cases}
\end{equation}

@@

We will treat the case when the branch is in the partner state ($s1=p$) in a similar way:
\begin{equation}
\begin{split}
p_{top=p}^{(h)}(t, t_r) &\approx 
\begin{cases}
\frac{p_{pb}^{(h)}(t) -e^{-(\lambda + \psi)(t_{h} - t)}}{U(t_{h})}&\textit{\color{gray} if $t_h \leq t_r$}\\
p_{pb}^{(r)}(t)\frac{p_{pa}^{(h)}(t_r) -e^{-(\lambda + \phi)(t_{h} - t_r)}}{U(t_{h})}&\textit{\color{gray} if $t \leq t_r \leq t_h$}\\
\frac{p_{pa}^{(h)}(t) -e^{-(\lambda + \phi)(t_{h} - t)}}{U(t_{h})}&\textit{\color{gray} if $t > t_r$}\\
\end{cases}\\
&\textit{\color{gray}where $t_r$ is the first notification time of the hidden partner.}\\
\end{split}
\end{equation}

Note that we wrote down equation~(\ref{eq:p_mixed}) as if the start of the hidden partner tree $t_h$ was known. In practice, it is not the case as we need to integrate over all the possible times between the branch start $t_{parent(i)}$ and its end $t_i)$. To simplify the calculations, we will take the time in the middle of this interval instead:

@@

Finally, we will treat the mixed case with two hidden partners using similar principals:
\begin{equation}
\begin{split}
p_{m(p,-,n)}^{(i)}(t,t_r) &\approx p_{top=p}^{(h1)}(t, t_r)U_p(t_{h1},t_{r}) p_{top=-}^{(h2)}(t_{h1}, t_i)U_p(t_{h2},t_{i}) p_n^{(i)}(t_{h2})\textit{\color{gray},}\\
&\textit{\color{gray}where $t_{h1} \approx t + \frac{1}{3}(t_i - t)$,}\\
&\textit{\color{gray}~~~~~~~~$t_{h2} \approx t + \frac{2}{3}(t_i - t)$,}\\
&\textit{\color{gray}~~~~~~~~$t_{r}$ is the earliest notification time of the partner }\\
&\textit{\color{gray}~~~~~~~~~~~~~~~~~~hidden at the top of the branch.}\\
\end{split}\label{eq:p_mixed_mixed}
\end{equation}

@@

%0 \textit{\color{gray} ~if $t_r < t_j$}\\
%\textit{\color{gray} (not possible as once notified partners do not transmit)}
%\\
%\\
p_{pa}^{(i)}(t_j)\phi(1 - \upsilon) \textit{\color{gray} ~if $i$ got notified at time $t_r < t_j$ and did not notify}\\
\textit{\color{gray} (notified partner branch evolution followed by notification-provoked sampling and no notification)}
~~~+p_{m(pa,-)}^{(i)}(t_j,t_r)\psi\rho(1-\upsilon) \textit{\color{gray} ~i.e. branch $i$ contains a hidden partner notified by $r$ at $t_r < t_j$, }\\
 \textit{\color{gray} ~~~(mixed partner-standard branch evolution followed by sampling and no notification)}\\
 ~~~+p_{m(pa,n)}^{(i)}(t_j,t_r)\psi\rho\upsilon \textit{\color{gray} ~i.e. branch $i$ contains a hidden partner notified both by $r$ and $i$}\\
 \textit{\color{gray} ~~~(mixed partner-notifier branch evolution followed by sampling and notification)}
~~~+p_{m(pa,-,n)}^{(i)}(t_j,t_r)\psi\rho\upsilon  \textit{\color{gray} ~i.e. branch $i$ contains a hidden partner notified by $r$ at $t_r < t_j$, }\\
\textit{\color{gray} ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~$i$ notified another hidden partner}\\
 \textit{\color{gray} ~~~(mixed partner-standard-notifier branch evolution followed by sampling and notification)}

@@

\textit{\color{gray} (partner-before-notification branch evolution followed by standard sampling and no notification)}

@@

~~~+p_{m(pb,-)}^{(i)}(t_j,t_r)\psi\rho(1-\upsilon) \textit{\color{gray} ~i.e. branch $i$ contains a hidden partner notified by $r$ at $t_r > t_i$, }\\

@@

 \textit{\color{gray} ~~~(mixed partner-standard branch evolution followed by sampling and no notification)}\\
 ~~~+p_{m(pb,n)}^{(i)}(t_j,t_i)\psi\rho\upsilon \textit{\color{gray} ~i.e. branch $i$ contains a hidden partner notified both by $r$ and $i$}\\
 \textit{\color{gray} ~~~(mixed partner-notifier branch evolution followed by sampling and notification)}

@@

~~~+p_{m(pb,-,n)}^{(i)}(t_j,t_r)\psi\rho\upsilon  \textit{\color{gray} ~i.e. branch $i$ contains a hidden partner notified by $r$ at $t_r > t_i$, }\\

@@

 \textit{\color{gray} ~~~(mixed partner-standard-notifier branch evolution followed by sampling and notification)}

@@

\textit{\color{gray} (partner-before-notification branch evolution followed by} \\

@@

~~~+p_{m(p,-)}^{(i)}(t_j,t_r)\psi\rho(1-\upsilon) \textit{\color{gray} ~i.e. branch $i$ contains a hidden partner notified by $r$ at $t_j \leq t_r \leq t_i$, }\\

@@

 \textit{\color{gray} ~~~(mixed partner-standard branch evolution followed by sampling and no notification)}\\
 ~~~+p_{m(p,n)}^{(i)}(t_j,t_r)\psi\rho\upsilon \textit{\color{gray} ~i.e. branch $i$ contains a hidden partner notified both by $r$ and $i$}\\
 \textit{\color{gray} ~~~(mixed partner-notifier branch evolution followed by sampling and notification)}

@@

~~~+p_{m(p,-,n)}^{(i)}(t_j,t_r)\psi\rho\upsilon  \textit{\color{gray} ~i.e. branch $i$ contains a hidden partner notified by $r$ at $t_j \leq t_r \leq t_i$, }\\

@@

l^{(n,n)}(i,r) &= \begin{cases}
p_b^{(i)}(t_j)\phi\upsilon \textit{\color{gray} ~if $i$ got notified at time $t_r < t_j$ and notified their most recent patner}\\
\textit{\color{gray} (notified notifier branch evolution followed by notification-provoked sampling and  notification)}\\

@@

p_n^{(r)}(t_j) p_b^{(i)}(t_r) \phi \upsilon \textit{\color{gray} ~if $i$ got notified at time $t_j \leq t_r \leq t_i$ and notified}\\
\textit{\color{gray} (notifier branch evolution followed by notified notifier branch evolution,} \\

@@

In the other case the internal node $i$'s branch starts as a partner who will eventually get notified by a tip $r$.
There are two possibilities: either the node $i$ corresponds to the partner or the partner is hidden along the $i$'s branch and $i$ corresponds to someone within $i$'s recipient subtree (e.g., if $i=C$ in Fig.~\ref{fig:pn-branches}(2a)). In the first case, the $i$'s branch evolution is represented by the partner branch evolution $p_{p}^{(i)}(t_j, t_r)$:

\begin{equation}
\scriptsize
p_{p}^{(i)}(t_j, t_r) =
\begin{cases}
p_{pa}^{(i)}(t_j) &\textit{\color{gray} if $t_r < t_j$,}\\
p_{pb}^{(i)}(t_j) &\textit{\color{gray} if $t_r > t_i$,}\\
p_{pb}^{(r)}(t_j)p_{pa}^{(i)}(t_r) &\textit{\color{gray} if $t_j \leq t_r \leq t_i$.}\\
\end{cases}\label{eq:p-p}
\end{equation}
This branch has a transmission event at the end of it, where the donor is known (probability density of $\lambda$): The donor corresponds to the subtree containing the tip representing the sampling of the individual $i$.

In the second case, the $i$'s branch evolution is represented by $p_{m(p,-)}^{(i)}(t_j,t_r)$ with a transmission event at the end of it, where the donor is unknown (probability density of $2\lambda$): any of the subtrees can be the donor.
 %Moreover, this recipient's branch cannot be notified by our partner (as in our model notified partners do not notify further):

Again, we will consider three possibilities for $i$'s child nodes: (2.2.1) when both $i_0$ and $i_1$ are internal nodes; (2.2.2) when both $i_0$ and $i_1$ are tips; and (2.2.3) when one of them is an internal node and the other one is a tip.

%In any of these scenarios, the notification must have happened after the partner tip branch start (as one of the model assumptions is that partners do not transmit after notification). Hence:
%\begin{equation}
%\scriptsize
%l^{(n)}(i, r) = 0 \textit{\color{gray}~if $t_i > t_r$} \\
%\label{eq:lp-zero}
%\end{equation}

@@

\begin{equation}
\scriptsize
\begin{split}
l^{(n)}(i, r) = &p_{p}^{(i)}(t_j, t_r) \lambda \cdot
\Big(~l^{(n)}(i_0,r)l^{(\not{n})}(i_1) \textit{\color{gray}~$\leftarrow$ $i_0$ is notified by $r$} \\
& ~~~~~~~~~~~~~~~~~~~+ l^{(\not{n})}(i_0)l^{(n)}(i_1,r)~\Big) \textit{\color{gray}~$\leftarrow$ $i_1$ is notified by $r$} \\
& + p_{m(p,-)}^{(i)}(t_j,t_r) 2 \lambda l^{(\not{n})}(i_0)l^{(\not{n})}(i_1) \textit{\color{gray}~$\leftarrow$ hidden partner}\\
 \end{split}
\label{eq:lp-int-int}
\end{equation}

@@

\begin{equation}
\scriptsize
\begin{split}
l^{(n)}(i, r) = &p_{p}^{(i)}(t_j,t_r) \lambda \cdot
\Big(~l^{(n,\not{n})}(i_0,first(r,i_1))l^{(\not{n},n)}(i_1) \textit{\color{gray}~$\leftarrow$ $r$ and $i_1$ notified $i_0$, $i_0$ did not notify} \\
&~~~~~~~~~~~~~~~~~~~+ l^{(n,n)}(i_0,first(r,i_1))l^{(n,n)}(i_1, i_0) \textit{\color{gray}~$\leftarrow$ $r$ and $i_1$ notified $i_0$, $i_0$ notified $i_1$} \\
&~~~~~~~~~~~~~~~~~~~+l^{(\not{n},n)}(i_0)l^{(n,\not{n})}(i_1,first(r,i_0)) \textit{\color{gray}~$\leftarrow$ $r$ and $i_0$ notified $i_1$, $i_1$ did not notify} \\
&~~~~~~~~~~~~~~~~~~~+ l^{(n,n)}(i_0, i_1)l^{(n,n)}(i_1,first(r,i_0)) \textit{\color{gray}~$\leftarrow$ $r$ and $i_0$ notified $i_1$, $i_1$ notified $i_0$} \\
&~~~~~~~~~~~~~~~~~~~+l^{(n,n)}(i_0,r)l^{(n,\not{n})}(i_1, i_0) \textit{\color{gray}~$\leftarrow$ $r$ notified $i_0$, $i_0$ notified $i_1$, $i_1$ did not notify} \\
&~~~~~~~~~~~~~~~~~~~+ l^{(n,\not{n})}(i_0, i_1)l^{(n,n)}(i_1,r) \textit{\color{gray}~$\leftarrow$ $r$ notified $i_1$, $i_1$ notified $i_0$, $i_0$ did not notify} \\
&~~~~~~~~~~~~~~~~~~~+l^{(n,\not{n})}(i_0,r)l^{(\not{n},\not{n})}(i_1) \textit{\color{gray}~$\leftarrow$ $r$ notified $i_0$, $i_0$ and $i_1$ did not notify} \\
&~~~~~~~~~~~~~~~~~~~+ l^{(\not{n},\not{n})}(i_0)l^{(n,\not{n})}(i_1,r)~\Big) \textit{\color{gray}~$\leftarrow$ $r$ notified $i_1$, $i_0$ and $i_1$ did not notify} 
\\
&+ p_{m(p,-)}^{(i)}(t_j,t_r) 2 \lambda \cdot
\Big(~l^{(n,\not{n})}(i_0,i_1)l^{(\not{n},n)}(i_1) \textit{\color{gray}~$\leftarrow$ $r$ notified a hidden partner,} \\
&\textit{\color{gray}~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~$i_1$ notified $i_0$, $i_0$ did not notify} \\
&~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~+ l^{(n,n)}(i_0,i_1)l^{(n,n)}(i_1, i_0)\textit{\color{gray}~$\leftarrow$ $r$ notified a hidden partner,} \\
&\textit{\color{gray}~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~$i_1$ notified $i_0$, $i_0$ notified $i_1$} \\
&~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~+l^{(\not{n},n)}(i_0)l^{(n,\not{n})}(i_1,i_0) \textit{\color{gray}~$\leftarrow$ $r$ notified a hidden partner,} \\
&\textit{\color{gray}~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~$i_0$ notified $i_1$, $i_1$ did not notify} \\
&~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~+l^{(\not{n},\not{n})}(i_0)l^{(\not{n},\not{n})}(i_1)~\Big) \textit{\color{gray}~$\leftarrow$ $r$ notified a hidden partner,} \\
&\textit{\color{gray}~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~$i_0$ and $i_1$ did not notify} \\
\\
&\textit{\color{gray}~where } \color{gray}first(a,b) = 
\begin{cases}
a \textit{~if $t_a \leq t_b$}\\
b \textit{~if $t_a > t_b$.}
\end{cases} 
 \end{split}
\label{eq:lu-tip-tip}
\end{equation}

@@

\begin{equation}
\scriptsize
\begin{split}
l^{(n)}(i, r) = &p_{p}^{(i)}(t_j,t_r) \lambda \cdot
\Big(~l^{(n,n)}(i_0,r)l^{(n)}(i_1,i_0) \textit{\color{gray}~$\leftarrow$ $r$ notified $i_0$, $i_0$  notified $i_1$} \\
&~~~~~~~~~~~~~~~~~~~+ l^{(n,\not{n})}(i_0,r)l^{(\not{n})}(i_1) \textit{\color{gray}~$\leftarrow$  $r$ notified $i_0$, $i_0$ did not notify} \\
&~~~~~~~~~~~~~~~~~~~+l^{(\not{n}, n)}(i_0)l^{(n)}(i_1,first(r,i_0)) \textit{\color{gray}~$\leftarrow$ both $r$ and $i_0$ notified $i_1$} \\
&~~~~~~~~~~~~~~~~~~~+ l^{(\not{n},\not{n})}(i_0)l^{(n)}(i_1,r)~\Big) \textit{\color{gray}~$\leftarrow$ $r$ notified $i_1$, $i_0$ did not notify} 
\\
&+ p_{m(p,-)}^{(i)}(t_j,t_r) 2\lambda \cdot
\Big(~l^{(\not{n},n)}(i_0)l^{(n)}(i_1,i_0) \textit{\color{gray}~$\leftarrow$ $r$ notified a hidden partner,} \\
&\textit{\color{gray}~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~$i_0$ notified $i_1$} \\
&~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~+ l^{(\not{n},\not{n})}(i_0)l^{(\not{n})}(i_1)~\Big) \textit{\color{gray}~$\leftarrow$ $r$ notified a hidden partner,} \\
&\textit{\color{gray}~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ $i_0$ did not notify}\\
\textit{\color{gray}~where } &\color{gray}first(a,b) = 
\begin{cases}
a \textit{~if $t_a \leq t_b$}\\
b \textit{~if $t_a > t_b$.}
\end{cases} 
 \end{split}
\label{eq:lp-tip-int}
\end{equation}
